---
title: "MaleHistology_Acini_Analysis"
author: "Samuel Gurr"
date: "December 16, 2020"
output: html_document
tags: geoduck pH reproductive development histology
---

# Load libraries
```{r}
library(dplyr)          # Version 0.7.6, Packaged: 2018-06-27, Depends: R (>= 3.1.2)Imports: assertthat (>= 0.2.0), bindrcpp (>= 0.2.0.9000), glue (>=1.1.1), magrittr (>= 1.5), methods, pkgconfig (>= 2.0.1), R6(>= 2.2.2), Rcpp (>= 0.12.15), rlang (>= 0.2.0), tibble (>=1.3.1), tidyselect (>= 0.2.3), utils
library(ggplot2)        # Version 2.2.1, Packaged: 2016-12-30, Depends: R (>= 3.1)Imports: digest, grid, gtable (>= 0.1.1), MASS, plyr (>= 1.7.1),reshape2, scales (>= 0.4.1), stats, tibble, lazyeval
library(tidyr)
library(ggpubr)
library(car)
library(gridExtra)
```

# set wd and Load data
```{r}
setwd("C:/Users/samjg/Documents/Github_Repositories/Pgenerosa_histology/RAnalysis/Data") #set working
staging<-read.csv("Staging/Hist_Staging_Kaitlyn.csv", header=T, sep=",", na.string="NA", as.is=T) 
Male_hist<-read.csv("Master_summary_male_acini.csv", header=T, sep=",", na.string="NA", as.is=T) 
head(staging)
head(Male_hist)
```

# Quantiative criteria 

```{r}

Male_hist_settings <- Male_hist %>% 
  dplyr::select(c('type', 'hue_range', 'saturation_range', 'brightness_range')) %>% 
  na.omit() %>% 
  dplyr::mutate(hue_start = as.numeric(sub("\\:.*", "", hue_range))) %>% 
  dplyr::mutate(hue_end = as.numeric(sub(".*:", "", hue_range))) %>% 
  dplyr::mutate(saturation_start = as.numeric(sub("\\:.*", "", saturation_range))) %>% 
  dplyr::mutate(saturation_end = as.numeric(sub(".*:", "", saturation_range))) %>% 
  dplyr::mutate(brightness_start = as.numeric(sub("\\:.*", "", brightness_range))) %>% 
  dplyr::mutate(brightness_end = as.numeric(sub(".*:", "", brightness_range))) %>% 
  dplyr::group_by(type)  %>%
  dplyr::summarize(hue_startMEAN = mean(hue_start, na.rm=TRUE),
                   hue_endMEAN   = mean(hue_end, na.rm=TRUE),
                   saturation_startMEAN = mean(saturation_start, na.rm=TRUE),
                   saturation_endMEAN   = mean(saturation_end, na.rm=TRUE),
                   brightness_startMEAN = mean(brightness_start, na.rm=TRUE),
                   brightness_endMEAN   = mean(brightness_end, na.rm=TRUE))

Male_hist_settings

```


# PIVOT THE TABLE TO CALC PROPORTION VALUES RELATIVE TO TOTAL AREA

```{r}
Male_hist_2 <- Male_hist %>% dplyr::select(-c('Meas_num','Label','hue_range','saturation_range','brightness_range')) %>% 
  tidyr::pivot_wider(names_from=type, values_from=Area)
Male_hist_2$perc_zoa <- (Male_hist_2$zoa/Male_hist_2$total_area)*100 # percent area of spermatozoa
Male_hist_2$perc_cytes <- ((Male_hist_2$cytes_zoa - Male_hist_2$zoa)/Male_hist_2$total_area)*100 # percent area of spermatocytes
Male_hist_2$perc_lumen <- (Male_hist_2$lumen/Male_hist_2$total_area)*100 # percent area of lumen
Male_hist_2$cytes <- (Male_hist_2$cytes_zoa - Male_hist_2$zoa) # area of spermatocytes
head(Male_hist_2)
```


# Average proprotions (persenct area) of each feature 
```{r}
Mean_Proportion_Table <- Male_hist_2 %>% 
  dplyr::select(c('Date', 'Treatment', 'perc_zoa','perc_cytes', 'perc_lumen')) %>% 
  na.omit() %>%
  dplyr::group_by(Date, Treatment) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())))
# Date      Treat   zoaMEAN    zoaSTE   perc_cytes_se
# 20190123	Ambient	8.995023   1.503110
# 20190123	Low	    11.272194  1.171135
# 20190221	Ambient	21.550115  2.871774
# 20190221	Low	    18.410939  1.872914	
```


#   TWO WAY ANOVAs (proprtion.hist.trait ~ treatment ? date)
```{r}
Male_hist_2$Date <- as.character(Male_hist_2$Date) # make Date a character variable!
# PERC ZOA TEST
PERCZoa_2factorialAOV <- aov(perc_zoa ~ Treatment * Date, data = Male_hist_2)
summary(PERCZoa_2factorialAOV) # date is a sig effect in raw data (not transformed!)
TukeyHSD(PERCZoa_2factorialAOV, which = "Date") # 02/21 > 01/23; increased with time
plot(PERCZoa_2factorialAOV, 1) # TEST ASSUMPTIONS     # 1. Homogeneity of variances (Levenes test) of the model data
leveneTest(perc_zoa ~ Treatment * Date, data = Male_hist_2) # PASSES
plot(PERCZoa_2factorialAOV, 2)  # TEST ASSUMPTIONS    # 2. Normality - QQ plot (quantile quantile) of the model residuals
PERCZoa_aov_residuals <- residuals(object = PERCZoa_2factorialAOV) # Extract the residuals
shapiro.test(x = PERCZoa_aov_residuals) # Run Shapiro-Wilk test; NOT NORMAL
hist(Male_hist_2$perc_zoa)
dev.off()
```
```{r}
# PERC LUMEN 
PERClumen_2factorialAOV <- aov(perc_lumen ~ Treatment * Date, data = Male_hist_2)
summary(PERClumen_2factorialAOV) # treatment and data are sig diff in raw data (not transformed!)
TukeyHSD(PERClumen_2factorialAOV, which = "Date") 
plot(PERClumen_2factorialAOV, 1) # TEST ASSUMPTIONS     # 1. Homogeneity of variances (Levenes test) of the model data
leveneTest(perc_lumen ~ Treatment * Date, data = Male_hist_2) # PASSES
plot(PERClumen_2factorialAOV, 2)  # TEST ASSUMPTIONS    # 2. Normality - QQ plot (quantile quantile) of the model residuals
PERClumen_aov_residuals <- residuals(object = PERClumen_2factorialAOV) # Extract the residuals
shapiro.test(x = PERClumen_aov_residuals) # Run Shapiro-Wilk test; NOT NORMAL
hist(Male_hist_2$perc_lumen)
dev.off()
```

```{r}
# PERC CYTES 
PERCcytes_2factorialAOV <- aov(perc_cytes ~ Treatment * Date, data = Male_hist_2)
summary(PERCcytes_2factorialAOV) # Date has a marginal diff (not transformed!)
TukeyHSD(PERCcytes_2factorialAOV, which = "Date") # 02/21 > 01/23; increased with time
plot(PERCcytes_2factorialAOV, 1) # TEST ASSUMPTIONS     # 1. Homogeneity of variances (Levenes test) of the model data
leveneTest(perc_cytes ~ Treatment * Date, data = Male_hist_2) # PASSED homogeneity of variance
plot(PERCcytes_2factorialAOV, 2)  # TEST ASSUMPTIONS    # 2. Normality - QQ plot (quantile quantile) of the model residuals
PERCcytes_aov_residuals <- residuals(object = PERCcytes_2factorialAOV) # Extract the residuals
shapiro.test(x = PERCcytes_aov_residuals) # Run Shapiro-Wilk test;  NORMAL
dev.off()
```

# BOXPLOTS RAW AREA DATA

```{r}
Male_hist_2$Treatment <- ordered(Male_hist_2$Treatment, levels = c("Ambient", "Low"))
cb_Colors <- c("#4E84C4", "#D16103")
# Zoa
zoa_plot <- ggplot(Male_hist_2, aes(x = Date, y = zoa, colour = Treatment)) +
  theme_bw() +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# lumen
lumen_plot <- ggplot(Male_hist_2, aes(x = Date, y = lumen, colour = Treatment)) +
  theme_bw() +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# cytes
cytes_plot <- ggplot(Male_hist_2, aes(x = Date, y = cytes, colour = Treatment)) +
  theme_bw() +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# grid plots
ALL_PLOTS <- grid.arrange(zoa_plot, lumen_plot, cytes_plot, ncol =3, nrow = 1)
ALL_PLOTS # view plot

```

# BOXPLOTS PROPORTION DATA

```{r}
# Zoa
zoa_plot_PERC <- ggplot(Male_hist_2, aes(x = Date, y = perc_zoa, colour = Treatment)) +
  theme_bw() +
  ylim(0,100) +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# lumen
lumen_plot_PERC <- ggplot(Male_hist_2, aes(x = Date, y = perc_lumen, colour = Treatment)) +
  theme_bw() +
  ylim(0,100) +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# cytes
cytes_plot_PERC <- ggplot(Male_hist_2, aes(x = Date, y = perc_cytes, colour = Treatment)) +
  theme_bw() +
  ylim(0,100) +
  geom_boxplot(outlier.size = 1, position=position_dodge(0.9), fill = "white") +
  geom_point(pch = 21, size = 3, fill = "white", position = position_jitterdodge(0.2)) +
  scale_color_manual(values=cb_Colors)
# grid plots
ALL_PLOTS_PERC <- grid.arrange(zoa_plot_PERC, lumen_plot_PERC, cytes_plot_PERC, ncol =3, nrow = 1)
ALL_PLOTS_PERC # view plot
```

# BOXPLOTS WITH STAGING 
```{r}
Male_hist_2$Geoduck_ID <-Male_hist_2$ID
staging_coltrim <- staging %>% dplyr::select(c('Geoduck_ID','Geoduck_ID','Stage_ID','Staging_number'))
MaleHistStage_merge <- merge(Male_hist_2,staging_coltrim,by="Geoduck_ID")

MaleHistStage_221 <- MaleHistStage_merge %>% dplyr::filter(Date %in% '20190221')
MaleHistStage_123 <- MaleHistStage_merge %>% dplyr::filter(Date %in% '20190123')

MaleHistStage_low <- MaleHistStage_merge %>% dplyr::filter(Treatment %in% 'Low')
MaleHistStage_ambient <- MaleHistStage_merge %>% dplyr::filter(Treatment %in% 'Ambient')

PERClumen_staging <- ggplot(MaleHistStage_merge, aes(factor(Staging_number), perc_lumen, fill= Treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#00BFC4","#F8766D")) +
  geom_point(size = 3, position = position_jitterdodge(jitter.width = 0.5))+
  theme_classic() +
  labs(y=expression("Lumen (% area total acini)"), x=expression("Staging ID")) +
  geom_smooth(method = "lm", se=FALSE, color="black", aes(group=Treatment)) +
  theme(legend.position = "none")
PERClumen_staging2 <-PERClumen_staging + theme(text = element_text(size = 15))# view plot

PERCzoa_staging <- ggplot(MaleHistStage_merge, aes(factor(Staging_number), perc_zoa, fill= Treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#00BFC4","#F8766D")) +
  geom_point(size = 3, position = position_jitterdodge(jitter.width = 0.5))+
  theme_classic() +
  labs(y=expression("Spermatozoa (% area total acini)"), x=expression("Staging ID")) +
  geom_smooth(method = "lm", se=FALSE, color="black", aes(group=Treatment)) +
  theme(legend.position = "none")
PERCzoa_staging2 <- PERCzoa_staging + theme(text = element_text(size = 15))# view plot

PERCcytes_staging <- ggplot(MaleHistStage_merge, aes(factor(Staging_number), perc_cytes, fill= Treatment)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#00BFC4","#F8766D")) +
  geom_point(size = 3, position = position_jitterdodge(jitter.width = 0.5))+
  theme_classic() +
  labs(y=expression("Spermatocytes (% area total acini)"), x=expression("Staging ID")) +
  geom_smooth(method = "lm", se=FALSE, color="black", aes(group=Treatment)) +
  theme(legend.position = "none")
PERCcytes_staging2 <-PERCcytes_staging + theme(text = element_text(size = 15))  # view plot

# grid plots
staging_hist_plots <- grid.arrange(PERClumen_staging2, PERCzoa_staging2, PERCcytes_staging2, ncol =3, nrow = 1)
staging_hist_plots # view plot
```


# STACKED PROPORTIONS BAR CHART
```{r}

# call subset for plotting and pivot longer with tidyr
colnames(Male_hist_2)
Male_hist_plots <- Male_hist_2[,c(1:3,9:11)]
Male_hist_plots_long <- Male_hist_plots %>% 
  tidyr::pivot_longer(cols = c(4:6), names_to='prop_metric', values_to='value')
# change labels for plotting with gsub
Male_hist_plots_long$Date <- gsub("20190123", "72",Male_hist_plots_long$Date)
Male_hist_plots_long$Date <- gsub("20190221", "93 + 8 day recovery",Male_hist_plots_long$Date)
Male_hist_plots_long$prop_metric <- gsub("perc_cytes", "spermatocytes",Male_hist_plots_long$prop_metric)
Male_hist_plots_long$prop_metric <- gsub("perc_zoa", "spermatozoa",Male_hist_plots_long$prop_metric)
Male_hist_plots_long$prop_metric <- gsub("perc_lumen", "lumen",Male_hist_plots_long$prop_metric)
#convert metric to ordered factor
Male_hist_plots_long_ordered <- Male_hist_plots_long[order(Male_hist_plots_long$prop_metric ),]
Male_hist_plots_long_ordered$prop_metric <- factor(Male_hist_plots_long_ordered$prop_metric, levels = rev(c(rep("lumen",1), rep("spermatocytes",1), rep("spermatozoa",1))))
Male_hist_plots_long_ordered_2 <- Male_hist_plots_long_ordered%>%
  group_by(Date, Treatment, prop_metric) %>% 
  dplyr::summarise(mean = mean(value), sd =sd(value)) %>% 
  mutate(y_pos = cumsum(mean))
# stacked proportion plot
pd <- position_dodge2(width = 0.2)
Male_hist_plots_long_ordered_2$prop_metric <- as.character(Male_hist_plots_long_ordered_2$prop_metric)

STACKED_barplot <- ggplot(Male_hist_plots_long_ordered_2, aes(x = Treatment, y = mean, fill = Treatment, alpha = prop_metric)) + 
  geom_bar(stat = "identity", width=0.7) + 
  scale_fill_manual(values = c("#00BFC4","#F8766D")) +
  geom_errorbar(aes(ymax = y_pos + sd, ymin=y_pos - sd), stat = "identity", width = 0.1, alpha = 0.7, position = pd) + 
  facet_wrap(~Date)+ 
  scale_alpha_manual(values=c(seq(0.3,1, length.out = 3))) + 
  theme_bw() + 
  scale_y_continuous(breaks = seq(from = 0, to = 100, by = 10)) + 
  ylim(0,100) +
  ylab("mean proportion of total acini area (%)") + 
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank()) 
STACKED_barplot

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
