---
title: "pH Female Broodstock Histology Analysis"
author: "Shelly Trigg"
date: "10/07/2020"
output: html_document
tags: geoduck pH reproductive development histology
editor_options: 
  chunk_output_type: console
---

Load libraries
```{r}
library(ggplot2)
library(ggpubr)
library(rstatix)
library(dplyr)
library(broom)
library(lme4)
library(kableExtra)
library(tidyr)
library(ggpmisc)
```

Read in data
```{r}
female_data <- read.csv("../../data/histology/Female_Gonad/20200820_Female_Gonad_RESULTS.csv", stringsAsFactors = F,colClasses = c(rep("character",4),rep("numeric",7)))


egg_size <- read.csv("../../data/histology/Female_Gonad/egg_size.csv", stringsAsFactors = F, colClasses = c("character","character","numeric"))

foll_size <- read.csv("../../data/histology/Female_Gonad/follicle_size.csv", stringsAsFactors = F, colClasses = c("character","character","numeric","numeric") )


male_quant <- read.csv("../../data/histology/Male_Gonad/Master_summary_male_acini.csv")

```


format male data
```{r}
#calculate proportions
male_quant_wd <- male_quant %>%dplyr::select(-c(Meas_num,Label,hue,saturation,brightness)) %>% pivot_wider(names_from=type, values_from=Area) %>% mutate(TotalAreaAscini = (lumen+zoa+(cytes_zoa - zoa))) %>% mutate(perc_zoa = zoa/TotalAreaAscini*100) %>% mutate(perc_cytes=(cytes_zoa - zoa)/TotalAreaAscini*100) %>% mutate(perc_lumen=lumen/TotalAreaAscini*100) %>% mutate(cytes = cytes_zoa - zoa)



#remove underscores from IDs
male_quant_wd$sample_ID <- gsub("_","",male_quant_wd$ID)
```

format egg size df
```{r}
#rename columns
colnames(egg_size) <- c("date", "sample_ID", "egg_size")

#merge egg size data with treatment info
egg_size_m <- merge(female_data[,c("date", "sample_ID","pH")], egg_size)

egg_size_m$date <- gsub("20190123", "72",egg_size_m$date)

egg_size_m$date <- gsub("20190221", "93 + 8 day recovery",egg_size_m$date)
```

format follicle size data
```{r}
foll_size <- foll_size[,c("Label", "Area")]

foll_size$date <- gsub("-.*","",foll_size$Label)

foll_size$sample_ID <- gsub("-10.*jpg","",foll_size$Label)

foll_size$sample_ID <- gsub(".*-","",foll_size$sample_ID)

foll_size <- foll_size[,c("Area", "sample_ID", "date")]

colnames(foll_size)[1] <- "foll_size"

#merge follicle size data with treatment info
foll_size_m <- merge(female_data[,c("date", "sample_ID","pH")], foll_size)

foll_size_m$date <- gsub("20190123", "72",foll_size_m$date)

foll_size_m$date <- gsub("20190221", "93 + 8 day recovery",foll_size_m$date)

```

plot egg size 
```{r}
#first check distribution

p <- ggplot(data = egg_size_m,aes(x = egg_size, color = pH, group = sample_ID)) + geom_density() + theme_bw() + labs(y = "density", x =expression(paste("egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"), axis.text.x = element_text(angle = 45)) + facet_wrap(~date)

#it's left skewed so try log transformation
q <- ggplot(data = egg_size_m,aes(x = log(egg_size,2), color = pH, group = sample_ID)) + geom_density() + theme_bw() + labs(y = "density", x = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"), axis.text.x = element_text(angle = 45)) + facet_wrap(~date)

#plot boxplots
r <- ggplot(data = egg_size_m,aes(x = date, y = egg_size, color = pH, group = interaction(pH,date, sample_ID))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, size=0.2,position = position_jitterdodge(jitter.width = 0.05)) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines")) + scale_color_manual(values = c("#00BFC4", "#F8766D"))

ggplot(data = egg_size_m, aes(x = pH, y = egg_size, color = pH)) + geom_boxplot(aes(group = pH), outlier.shape = NA) + geom_point(aes(group = interaction(pH, sample_ID)),pch = 16, size=0.2,position = position_jitterdodge(jitter.width = 0.05)) + geom_boxplot(outlier.shape = NA, fill = NA)+ theme_bw() + labs(x = "exposure time (days)", y = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"), legend.position = "bottom") + scale_color_manual(values = c("#00BFC4", "#F8766D")) + facet_wrap(~date)

#after log transformation
s <- ggplot(data = egg_size_m,aes(x = date, y = log(egg_size,2), color = pH, group = interaction(pH,date, sample_ID))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, size=0.2,position = position_jitterdodge(jitter.width = 0.05)) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"))

#plot egg size without transformation
ggpubr::ggarrange(p,r,labels = "AUTO", common.legend = T)


#plot eggs size after transformation
ggpubr::ggarrange(q,s,labels = "AUTO", common.legend = T)
```

Save plots
```{r}
jpeg("egg_size_dist.jpg", width = 12, height = 6, units = "in", res = 300)
ggpubr::ggarrange(p,r,labels = "AUTO", common.legend = T)
dev.off()

jpeg("log_egg_size_dist.jpg", width = 12, height = 6, units = "in", res = 300)
ggpubr::ggarrange(q,s,labels = "AUTO", common.legend = T)
dev.off()
```


Egg size stats
```{r}
tidy(TukeyHSD(aov(log(egg_size,2) ~ pH * date, data = egg_size_m))) %>%
  kbl(caption = "2-way ANOVA: log2 egg size ~ pH * time") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Save 2-way ANOVA results
```{r}
write.table(tidy(TukeyHSD(aov(log(egg_size,2) ~ pH * date, data = egg_size_m))),"2wayAOV_Tuk_egg_size.txt",sep = "\t", quote = F, row.names = F)
```


Egg size mixed model stats
```{r}
anova(lme(log(egg_size,2) ~ pH * date, random = ~1|sample_ID, data = egg_size_m)) %>%
  kbl(caption = "linear mixed model: log2 egg size ~ pH * time, random = sample_ID") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```


plot foll size 
```{r}
#first check distribution

p <- ggplot(data = foll_size_m,aes(x = foll_size, color = pH, group = sample_ID)) + geom_density() + theme_bw() + labs(y = "density", x =expression(paste("egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"), axis.text.x = element_text(angle = 45)) + facet_wrap(~date)

#it's left skewed so try log transformation
q <- ggplot(data = foll_size_m,aes(x = log(foll_size,2), color = pH, group = sample_ID)) + geom_density() + theme_bw() + labs(y = "density", x = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"), axis.text.x = element_text(angle = 45)) + facet_wrap(~date)

#plot boxplots
r <- ggplot(data = foll_size_m,aes(x = date, y = foll_size, color = pH, group = interaction(pH,date, sample_ID))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, size=1,position = position_jitterdodge(jitter.width = 0.05)) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste(log[2]," follicle size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"))

#after log transformation
s <- ggplot(data = foll_size_m,aes(x = date, y = log(foll_size,2), color = pH, group = interaction(pH,date, sample_ID))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, size=1,position = position_jitterdodge(jitter.width = 0.05)) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"))

#plot egg size without transformation
ggpubr::ggarrange(p,r,labels = "AUTO", common.legend = T)


#plot eggs size after transformation
ggpubr::ggarrange(q,s,labels = "AUTO", common.legend = T)
```

Save plots
```{r}
jpeg("foll_size_dist.jpg", width = 12, height = 6, units = "in", res = 300)
ggpubr::ggarrange(p,r,labels = "AUTO", common.legend = T)
dev.off()

jpeg("log_foll_size_dist.jpg", width = 12, height = 6, units = "in", res = 300)
ggpubr::ggarrange(q,s,labels = "AUTO", common.legend = T)
dev.off()
```


Tissue proportions analysis
```{r}
#calculate proportions

female_data$prop_egg <- (female_data$egg.area..based.on.ind.eggs. / female_data$total.area..um.) *100

female_data$prop_no_tissue <- (female_data$no.tissue.area / female_data$total.area..um.) *100

female_data$prop_cnx_tissue_indve <- (female_data$connective.tissue.area..egg.based.on.indiv./ female_data$total.area..um.)*100

#create column for relative # of eggs
female_data$number.of.eggs.per.um2 <- (female_data$number.of.eggs/female_data$total.area..um.)

#merge median egg size with female data
female_data <- merge(female_data, egg_size_m %>% group_by(date, pH, sample_ID) %>% summarise(med.egg.size = median(egg_size))  %>% ungroup() %>% dplyr::select(sample_ID, med.egg.size), by = "sample_ID")

female_data$pH_date <- paste(female_data$date, female_data$pH, sep = "_")

#convert female data to long format

female_data_STACKED <- tidyr::gather(female_data,"metric", "um", 5:(ncol(female_data)-1))

female_data_STACKED$um <- as.numeric(female_data_STACKED$um)




#reformat data and calculate mean and sd proportions
female_data_STACKED_props <- female_data_STACKED[grep("prop", female_data_STACKED$metric),]%>%group_by(date, pH,metric) %>% summarise(mean = mean(um), sd =sd(um)) %>% mutate(y_pos = cumsum(mean))

#convert metric to ordered factor
female_data_STACKED_props$metric <- factor(female_data_STACKED_props$metric, levels = rev(c("prop_cnx_tissue_indve","prop_egg", "prop_no_tissue")))


#chi sq to compare 20190123 amb vs. low
ct <- data.frame(female_data_STACKED_props[grep("20190123", female_data_STACKED_props$date),c("pH","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_a <- chisq.test(ct)
ct_a <- tidy(ct_a)
ct_a$comparison <- "ambient:72-low:72"


#chi sq to compare 20190123 amb vs. 20190221 amb
ct <- data.frame(female_data_STACKED_props[grep("ambient", female_data_STACKED_props$pH),c("date","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_b <- chisq.test(ct)
ct_b <- tidy(ct_b)
ct_b$comparison <- "ambient:72-ambient:93 + 8 day recovery"

#chi sq to compare 20190123 low vs. 20190221 low
ct <- data.frame(female_data_STACKED_props[grep("low", female_data_STACKED_props$pH),c("date","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_c <- chisq.test(ct)
ct_c <- tidy(ct_c)
ct_c$comparison <- "low:72-low:93 + 8 day recovery"



#chi sq to compare 20190221 amb vs. 20190221 low
ct <- data.frame(female_data_STACKED_props[grep("20190221", female_data_STACKED_props$date),c("pH","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_d <- chisq.test(ct)
ct_d <- tidy(ct_d)
ct_d$comparison <- "ambient:93 + 8 day recovery-low:93 + 8 day recovery"


#chi sq to compare 20190123 amb vs. 20190221 low
ct <- data.frame(female_data_STACKED_props[which(female_data_STACKED_props$date=="20190123" & female_data_STACKED_props$pH =="ambient" | female_data_STACKED_props$date=="20190221" & female_data_STACKED_props$pH =="low"),c("pH","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_e <- chisq.test(ct)
ct_e <- tidy(ct_e)
ct_e$comparison <- "ambient:72-low:93 + 8 day recovery"


#chi sq to compare 20190221 amb vs. 20190123 low
ct <- data.frame(female_data_STACKED_props[which(female_data_STACKED_props$date=="20190221" & female_data_STACKED_props$pH =="ambient" | female_data_STACKED_props$date=="20190123" & female_data_STACKED_props$pH =="low"),c("pH","mean","metric")]%>% pivot_wider(names_from = metric, values_from = mean))

rownames(ct) <- ct$pH

ct <- ct[,-1]

ct_f <- chisq.test(ct)
ct_f <- tidy(ct_f)
ct_f$comparison <- "ambient:93 + 8 day recovery-low:72"


ct <- rbind(ct_a, ct_b, ct_c, ct_d, ct_e, ct_f)

#perform FDR correction for multiple comparisons

ct$adj.p.value <- p.adjust(ct$p.value)

#reorder columns

ct <- ct[,c("comparison", "statistic", "p.value", "adj.p.value")]


ct %>%
  kbl(caption = "Chi square test: tissue proportions ~ pH") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


```{r, eval=F, "plot stacked bar of tissue areas"}
pd <- position_dodge2(width = 2)


jpeg("tissue_proportion_plots.jpg", width = 5, height = 5, units = "in", res = 300)
female_data_STACKED_props %>% mutate(date = ifelse(date =="20190123","72 d","8 d post")) %>% mutate(pH = ifelse(pH == "low","elevated pCO2","ambient pCO2")) %>% mutate(metric = gsub("prop_cnx_tissue_indve", "connective tissue", metric)) %>% mutate(metric = gsub("prop_egg", "eggs", metric)) %>% mutate(metric =gsub("prop_no_tissue", "follicles", metric)) %>% mutate(metric = factor(metric, levels = c("follicles", "eggs", "connective tissue")))%>% ggplot( aes(pH,mean,fill = metric)) + 
geom_bar(stat = "identity") + 
geom_errorbar(aes(ymax = y_pos + sd, ymin=y_pos - sd), stat = "identity", alpha = 0.7, width = 0,position = position_dodge(width = 0.3), color = "black") + facet_wrap(~date) + theme_bw() + scale_y_continuous(breaks = seq(from = 0, to = 100, by = 10)) + labs(x = "treatment",y ="percent total area (mean ± SD)", fill = "") + scale_fill_manual(values = gray.colors(n =3,0.4,0.8)) +theme(legend.position = "bottom") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(text = element_text(size = 12))   
dev.off()
```


add qualitative staging data
```{r}
#read in stage data
staging <- read_xlsx("../../data/histology/HistologyScores_STRIGG.xlsx")

#convert sample ID name to match other data
colnames(staging) <- gsub("Geoduck.ID", "sample_ID", colnames(staging))

#convert date to full date
staging <- staging %>% mutate(Date = ifelse(Date == "123", "20190123", "20190221"))

#combine stage data with quant scoring data
quant_qual <- merge(female_data_STACKED, staging, by = "sample_ID")

quant_qual_m <- merge(male_quant_wd,staging, by = c("sample_ID","Date", "Treatment"))

#replace NAs in grace's data with shelly's data
quant_qual_m$Stage_Crandall <- ifelse(quant_qual_m$Stage_Crandall=="NA",quant_qual_m$Stage_Trigg,quant_qual_m$Stage_Crandall)

#create PCO2 column
quant_qual_m <- quant_qual_m %>% mutate(pCO2 = factor(ifelse(Treatment == "Low","elevated", "ambient"), levels = c("elevated","ambient"))) 

#calculate means for each animal
quant_qual_m_summ <- quant_qual_m %>% group_by(sample_ID, pCO2, Stage_Crandall, Stage_Trigg,Sex,Date) %>% summarise(mean_perc_zoa = mean(perc_zoa), sd_perc_zoa = sd(perc_zoa), mean_perc_cytes = mean(perc_cytes), sd_perc_cytes = sd(perc_cytes), mean_perc_lumen = mean(perc_lumen), sd_perc_lumen = sd(perc_lumen))


#reformat Female data
quant_qual_long <- quant_qual %>% pivot_longer(cols = 13:14, names_to = "scorer", values_to = "qual_stage") %>% mutate(scorer = gsub("Stage_","", scorer)) %>% mutate(pCO2 = factor(ifelse(pH == "low","elevated", "ambient")))

#plot egg area x stage
jpeg("feature_area_by_staging_reg.jpg", width = 10, height = 8, units = "in", res = 300)

a <- quant_qual_long %>% filter(metric == "prop_egg") %>% filter(scorer == "Crandall") %>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(rr.label), "*\", \"*", after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "eggs (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_y_continuous(limits =c(0,25)) + scale_x_continuous(n.breaks =3) +  scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

#plot follicle area x stage
b <- quant_qual_long%>% filter(metric == "prop_no_tissue") %>% filter(scorer == "Crandall") %>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size = 0.5)+ geom_point(aes(shape = pCO2), size = 2)  + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)  + labs(x = "stage", y = "follicles (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


#plot conx tissue area x stage
c <- quant_qual_long %>% filter(metric == "prop_cnx_tissue_indve") %>% filter(scorer == "Crandall")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer)))  + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size =0.5)+ geom_point(aes(shape = pCO2),size = 2) + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8) + labs(x = "stage", y = "connective tissue (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


d <- quant_qual_m_summ %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_lumen)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "lumen (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

e <- quant_qual_m_summ  %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_zoa)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatozoa (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

f <- quant_qual_m_summ   %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_cytes)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "right", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatocytes (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

ggarrange(a,b,c,d,e,f, ncol = 3,nrow = 2, labels = "auto", common.legend = T, legend = "bottom")
dev.off()

#quant_qual_m_summ %>% filter(Date == "20190123") %>% ggplot(aes(x = as.numeric(Stage_Trigg), y = mean_perc_zoa)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(color = pCO2, shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatozoa (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(17,16)) + theme(legend.key=element_rect(fill="white"))



jpeg("egg_size_by_staging_reg.jpg", width = 8, height = 4, units = "in", res = 300)
g <- egg_size_m %>% mutate(pCO2 = ifelse(pH =="low", "elevated","ambient"))%>% mutate(date = ifelse(date == "72","72 d", "8 d post")) %>% mutate(sample_ID = factor(sample_ID, levels = c("045","050","052","13","17","006","028","034","053","057","059","060","14","6","7","8","009","015","025"))) %>% ggplot(aes(x = date, y = log(egg_size,2), color = pCO2, shape = pCO2,group = interaction(pCO2,date, sample_ID))) + geom_point(size=0.1,position = position_jitterdodge(jitter.width = 0.05))  + geom_boxplot(outlier.shape = NA, alpha = 0.4)+ theme_bw() + labs(shape = expression(paste(italic("p"),"CO", sep = "")[2]), color = expression(paste(italic("p"),"CO", sep = "")[2]),x = "sampling time point", y = expression(paste(log[2]," egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines")) + scale_shape_manual(values = c(1, 16)) + scale_color_manual(values = c("gray60", "gray30")) + theme(legend.position = "none")

h <- quant_qual_long %>% filter(metric == "mean.egg.size") %>% filter(scorer == "Crandall")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer)))  + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size =0.5)+ geom_point(aes(shape = pCO2),size = 2) + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 4) + labs(x = "stage", y = expression(paste("mean egg size (",mu,"m"^2,")", sep = "")), color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))

ggarrange(g,h, ncol = 2, labels = "auto", common.legend = F, legend = "top")
dev.off()







#plot egg area x stage
jpeg("feature_area_by_staging_reg_d72.jpg", width = 10, height = 8, units = "in", res = 300)

a <- quant_qual_long %>% filter(metric == "prop_egg") %>% filter(scorer == "Crandall") %>% filter(date =="20190123") %>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "eggs (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_y_continuous(limits =c(0,25)) + scale_x_continuous(n.breaks =3) +  scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

#plot follicle area x stage
b <- quant_qual_long%>% filter(metric == "prop_no_tissue") %>% filter(scorer == "Crandall")  %>% filter(date =="20190123")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size = 0.5)+ geom_point(aes(shape = pCO2), size = 2)  + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)  + labs(x = "stage", y = "follicles (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


#plot conx tissue area x stage
c <- quant_qual_long %>% filter(metric == "prop_cnx_tissue_indve") %>% filter(scorer == "Crandall") %>% filter(date =="20190123")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer)))  + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size =0.5)+ geom_point(aes(shape = pCO2),size = 2) + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8) + labs(x = "stage", y = "connective tissue (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


d <- quant_qual_m_summ%>% filter(Date =="20190123")%>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_lumen)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "lumen (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

e <- quant_qual_m_summ %>% filter(Date =="20190123") %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_zoa)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatozoa (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

f <- quant_qual_m_summ  %>% filter(Date =="20190123") %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_cytes)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "right", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatocytes (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

ggarrange(a,b,c,d,e,f, ncol = 3,nrow = 2, labels = "auto", common.legend = T, legend = "bottom")
dev.off()

jpeg("feature_area_by_staging_reg_8dPost.jpg", width = 10, height = 8, units = "in", res = 300)

a <- quant_qual_long %>% filter(metric == "prop_egg") %>% filter(scorer == "Crandall") %>% filter(date =="20190221") %>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "eggs (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_y_continuous(limits =c(0,25)) + scale_x_continuous(n.breaks =3) +  scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

#plot follicle area x stage
b <- quant_qual_long%>% filter(metric == "prop_no_tissue") %>% filter(scorer == "Crandall")  %>% filter(date =="20190221")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer))) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size = 0.5)+ geom_point(aes(shape = pCO2), size = 2)  + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)  + labs(x = "stage", y = "follicles (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


#plot conx tissue area x stage
c <- quant_qual_long %>% filter(metric == "prop_cnx_tissue_indve") %>% filter(scorer == "Crandall") %>% filter(date =="20190221")%>% ggplot(aes(x = as.numeric(qual_stage), y = um, group = interaction(sample_ID,scorer)))  + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80", size =0.5)+ geom_point(aes(shape = pCO2),size = 2) + stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8) + labs(x = "stage", y = "connective tissue (% total area)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + scale_x_continuous(n.breaks =3)+theme(legend.key=element_rect(fill="white"))


d <- quant_qual_m_summ%>% filter(Date =="20190221")%>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_lumen)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "lumen (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

e <- quant_qual_m_summ %>% filter(Date =="20190221") %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_zoa)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "left", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatozoa (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

f <- quant_qual_m_summ  %>% filter(Date =="20190221") %>% ggplot(aes(x = as.numeric(Stage_Crandall), y = mean_perc_cytes)) + geom_smooth(aes(group = Sex),method = "lm", color="black", fill = "gray80",size = 0.5)  + geom_point(aes(shape = pCO2),size = 2)+ stat_poly_eq(aes(group = Sex,label = paste(after_stat(adj.rr.label), "*\", \"*", after_stat(p.value.label),sep = "")), label.x = "right", label.y = "top",formula = y~x, parse = TRUE, size = 2.8)+ labs(x = "stage", y = "spermatocytes (% area total acini)", color =expression(paste(italic("p"),"CO", sep = "")[2]),shape =expression(paste(italic("p"),"CO", sep = "")[2])) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_shape_manual(values = c(1,16)) + theme(legend.key=element_rect(fill="white"))

ggarrange(a,b,c,d,e,f, ncol = 3,nrow = 2, labels = "auto", common.legend = T, legend = "bottom")
dev.off()
```


Write out Chi sq. test results
```{r}
write.table(ct,"chisq_tissue_prop.txt",sep = "\t", quote = F, row.names = F)
```

Generate stacked bars showing average tissue proportions
```{r}

#export plot with significance denoted
#plot stacked bar of tissue areas



#change labels for plotting
female_data_STACKED_props_labs <- female_data_STACKED_props

female_data_STACKED_props_labs$date <- gsub("20190123", "72 days of exposure",female_data_STACKED_props_labs$date)

female_data_STACKED_props_labs$date <- gsub("20190221", "93 days of exposure \n + 8 day recovery",female_data_STACKED_props_labs$date)

female_data_STACKED_props_labs$metric <- gsub("prop_no_tissue","follicle area", female_data_STACKED_props_labs$metric)

female_data_STACKED_props_labs$metric <- gsub("prop_cnx_tissue_indve","connective tissue area", female_data_STACKED_props_labs$metric)

female_data_STACKED_props_labs$metric <- gsub("prop_egg","egg area", female_data_STACKED_props_labs$metric)

#convert metric to ordered factor
female_data_STACKED_props_labs <- female_data_STACKED_props_labs[order(female_data_STACKED_props_labs$metric),]

female_data_STACKED_props_labs$metric <- factor(female_data_STACKED_props_labs$metric, levels = rev(c(rep("connective tissue area",1), rep("egg area",1), rep("follicle area",1))))

pd <- position_dodge2(width = 0.2)

ggplot(female_data_STACKED_props_labs, aes(pH,mean, fill = pH,alpha = metric)) + 
geom_bar(stat = "identity", color = "black") + 
geom_errorbar(aes(ymax = y_pos + sd, ymin=y_pos - sd), stat = "identity", width = 0.1, alpha = 0.7, position = pd) + facet_wrap(~date)+ scale_alpha_manual(values=c(seq(0.3,1, length.out = 3))) + theme_bw() + scale_y_continuous(breaks = seq(from = 0, to = 100, by = 10)) + ylab("mean proportion of total tissue area (%)") + theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank()) + scale_fill_manual(values = c("#00BFC4","#F8766D"))


jpeg("tissue_proportions_stacked_bar.jpg", width = 7, height = 6, units = "in", res = 300)
ggplot(female_data_STACKED_props_labs, aes(pH,mean, fill = pH,alpha = metric)) + 
geom_bar(stat = "identity", color = "black") + 
geom_errorbar(aes(ymax = y_pos + sd, ymin=y_pos - sd), stat = "identity", width = 0.1, alpha = 0.7, position = pd) + facet_wrap(~date)+ scale_alpha_manual(values=c(seq(0.3,1, length.out = 3))) + theme_bw() + scale_y_continuous(breaks = seq(from = 0, to = 100, by = 10)) + ylab("percent of total tissue area (mean ± s.d.)") + theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank()) + scale_fill_manual(values = c("#00BFC4", "#F8766D")) + ggtitle("Female quantitative histology scoring")
dev.off()


pdf("tissue_proportions_stacked_bar.pdf", width = 7, height = 6)
ggplot(female_data_STACKED_props_labs, aes(pH,mean, fill = pH,alpha = metric)) + 
geom_bar(stat = "identity", color = "black") + 
geom_errorbar(aes(ymax = y_pos + sd, ymin=y_pos - sd), stat = "identity", width = 0.1, alpha = 0.7, position = pd) + facet_wrap(~date)+ scale_alpha_manual(values=c(seq(0.3,1, length.out = 3))) + theme_bw() + scale_y_continuous(breaks = seq(from = 0, to = 100, by = 10)) + ylab("percent of total tissue area (mean ± s.d.)") + theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank()) + scale_fill_manual(values = c("#00BFC4", "#F8766D")) + ggtitle("Female quantitative histology scoring")
dev.off()

```



filter data for comparing
```{r, eval=F}
female_data_STACKED_filt <- female_data_STACKED[grep("prop|per|mean", female_data_STACKED$metric),]

```

Convert character to factors
```{r, eval = F}
female_data_STACKED_filt$pH <- as.factor(female_data_STACKED_filt$pH)

female_data_STACKED_filt$date <- as.factor(female_data_STACKED_filt$date)

female_data_STACKED_filt$metric <- as.factor(female_data_STACKED_filt$metric)



```

Plot each metric x treatment group
```{r, eval = F}

#plot january samples
#jpeg("20190123_boxplots.jpg", width = 6, height = 6, units = "in", res = 300)
a <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$date == "20190123"),],aes(x = pH, y = um, color = pH)) + geom_boxplot(outlier.shape = NA) + facet_wrap(~metric, scale = "free") + geom_jitter(alpha=0.8,shape =16,size = 2, position= position_jitter(0.1)) + theme_bw() + ggtitle("20190123_samples; ambient pH (n = 5) x low pH (n = 3)")
#dev.off()

#plot feb samples
#jpeg("20190221_boxplots.jpg", width = 6, height = 6, units = "in", res = 300)
b <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$date == "20190221"),],aes(x = pH, y = um, color = pH)) + geom_boxplot(outlier.shape = NA) + facet_wrap(~metric, scale = "free") + geom_jitter(alpha=0.8,shape =16,size = 2, position= position_jitter(0.1)) + theme_bw() + ggtitle("20190221_samples; ambient pH (n = 8) x low pH (n = 3)")
#dev.off()

jpeg("all_samples_boxplots.jpg", width = 12, height = 6, units = "in", res = 300)
ggpubr::ggarrange(a,b,labels = "AUTO", common.legend = T)
dev.off()

#plot all samples together
#jpeg("dates_combined_boxplots.jpg", width = 6, height = 6, units = "in", res = 300)
#ggplot(data = female_data_STACKED_filt,aes(x = pH, y = um)) +geom_boxplot(aes(color = pH), outlier.shape = NA) + facet_wrap(~metric, scale = "free") + geom_jitter(aes(color = pH),alpha = 0.8,shape =16, size = 2,position= position_jitter(0.2)) + theme_bw() + ggtitle("all samples; ambient pH (n = 13) x low pH (n = 6)")
#dev.off()
```

clarify pH treatment and dates
```{r, eval = F}
female_data_STACKED_filt$pH <- gsub("low", "low (6.8)", female_data_STACKED_filt$pH)

female_data_STACKED_filt$pH <- gsub("ambient", "ambient (7.8)", female_data_STACKED_filt$pH)

female_data_STACKED_filt$date <- gsub("20190123", "72",female_data_STACKED_filt$date)

female_data_STACKED_filt$date <- gsub("20190221", "93 + 8 day recovery",female_data_STACKED_filt$date)
```


plot egg size only 
```{r, eval = F}
c <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric == "mean.egg.size"),],aes(x = date, y = um, color = pH, group = interaction(pH,date))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, position = position_jitterdodge()) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste("mean egg size (",mu,"m"^2,")", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"),, axis.text.x = element_text(angle = 45))

d <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric == "number.of.eggs.per.um2"),],aes(x = date, y = um, color = pH, group = interaction(pH,date))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, position = position_jitterdodge()) + theme_bw() + labs(x = "exposure time (days)", y = expression(paste("number of eggs per ",mu,"m"^2," tissue", sep = "")))+ theme(plot.margin = unit(c(rep(1,4)), "lines"),, axis.text.x = element_text(angle = 45))

e <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric == "prop_cnx_tissue_indve"),],aes(x = date, y = um, color = pH, group = interaction(pH,date))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, position = position_jitterdodge()) + theme_bw() + labs(x = "exposure time (days)", y = "connective tissue\nproportion (%)")+ theme(plot.margin = unit(c(rep(1,4)), "lines"),, axis.text.x = element_text(angle = 45))

f <- ggplot(data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric == "prop_no_tissue"),],aes(x = date, y = um, color = pH, group = interaction(pH,date))) + geom_boxplot(outlier.shape = NA) + geom_point(pch = 16, position = position_jitterdodge()) + theme_bw() + labs(x = "exposure time (days)", y = "follicle area proportion (%)") + theme(plot.margin = unit(c(rep(1,4)), "lines"), axis.text.x = element_text(angle = 45))


jpeg("all_samples_boxplots_newlabs.jpg", width = 10, height = 3, units = "in", res = 300)

ggpubr::ggarrange(c,d,e,f, labels = "AUTO", vjust = 1,common.legend = T, nrow = 1,ncol = 4,legend = "bottom", font.label = c(size = 18), align = "hv")
dev.off()
```


Test for treatment and time effect
```{r, eval = F}
## egg size test---------------------------------
#egg size normality with shapiro test
female_data_STACKED_filt %>% tidyr::pivot_wider(names_from = metric,values_from = um ) %>% shapiro_test(mean.egg.size)
#  variable      statistic     p
#  <chr>             <dbl> <dbl>
#1 mean.egg.size     0.951 0.407

#levene test
female_data_STACKED_filt %>% tidyr::pivot_wider(names_from = metric,values_from = um ) %>% levene_test(mean.egg.size~pH)
# A tibble: 1 x 4
#    df1   df2 statistic     p
#  <int> <int>     <dbl> <dbl>
#1     1    17  0.000532 0.982



#try a two way ANOVA for effects of pH and time
egg_size_pHxTIME <- aov(um~date*pH, data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric=="mean.egg.size"),])

summary(egg_size_pHxTIME)
#            Df Sum Sq Mean Sq F value Pr(>F)  
#date         1  31.49  31.491   4.941  0.042 *
#pH           1   6.07   6.074   0.953  0.344  
#date:pH      1   4.97   4.974   0.780  0.391  
#Residuals   15  95.60   6.373

#try a 1way anova for time effect
egg_size_TIME  <- aov(um~date, data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric=="mean.egg.size"),])

summary(egg_size_TIME)

#          Df Sum Sq Mean Sq F value Pr(>F)  
#date         1  31.49  31.491    5.02 0.0387 *
#Residuals   17 106.65   6.273                 
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


## Date has an effect

#Does pH have an effect?
egg_size_pH  <- aov(um~pH, data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric=="mean.egg.size"),])

summary(egg_size_pH)
#            Df Sum Sq Mean Sq F value Pr(>F)
#pH           1   9.36   9.361   1.236  0.282
#Residuals   17 128.78   7.575     


#is there an effect of treatment at a specific time?
egg_size_jan_pH  <- aov(um~pH, data = female_data_STACKED_filt[which(female_data_STACKED_filt$metric=="mean.egg.size" & female_data_STACKED_filt$date == "20190123"),])

#yes there is 
summary(egg_size_jan_pH)
#            Df Sum Sq Mean Sq F value Pr(>F)  
#pH           1 10.963  10.963   8.825 0.0249 *
#Residuals    6  7.453   1.242                 
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#BUT NOT in the Feb group
#summary(egg_size_feb_pH)
#            Df Sum Sq Mean Sq F value Pr(>F)
#pH           1   0.08   0.085   0.009  0.928
#Residuals    9  88.14   9.794

```

Run wilcox test to see if follicle area is significantly different
```{r, eval =F}

wt <- female_data_STACKED_filt %>% group_by(metric) %>% wilcox_test(um ~ pH )


glm_res <- female_data_STACKED_filt %>% group_by(metric) %>% do(glm_models = glm(um ~ pH*date, data = .))

full.lmer <-  female_data_STACKED_filt %>% group_by(metric) %>% do(lmer_models = lmer(um ~ pH*date + (1|tank), data = ., REML = F))


red.lmer <-  female_data_STACKED_filt %>% group_by(metric) %>% do(lmer_models = lmer(um ~ 1 + (1|tank), data = ., REML = F))

for(i in 1:length(full.lmer$metric)){
  a <- anova(red.lmer$lmer_models[[i]], full.lmer$lmer_models[[i]])
  print(a)
}


aov_res_summ <- broom::tidy(aov_res)

wt <- wt %>% add_xy_position(x = "pH")

bxp <- ggboxplot(female_data_STACKED_filt, x = "pH", y = "um", color = "pH",facet.by = "metric", scales = "free", add = "jitter") 

jpeg("dates_combined_boxplots_pvals.jpg", width = 6, height = 6, units = "in", res = 300)
bxp + stat_pvalue_manual(wt,bracket.nudge.y = -2,label = "{p}") + scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()


wwt <- female_data_STACKED_filt %>% group_by(date,metric) %>% do(broom::tidy(wilcox.test(um ~ pH,data = . )))



#run stats on all data regardless of date
wt <- female_data_STACKED_filt %>% group_by(metric) %>% do(broom::tidy(wilcox.test(um ~ pH,data = . )))


#run stats on all data regardless of date
Tt <- female_data_STACKED_filt %>% group_by(metric) %>% do(broom::tidy(t.test(um ~ pH,data = . )))

aovt <- female_data_STACKED_filt %>% group_by(metric) %>% do(broom::tidy(aov(um ~ pH,data = . )))

glmt <- female_data_STACKED_filt %>% group_by(metric) %>% do(broom::tidy(glm(um ~ pH,data = . )))

```

Plot percent follicle area for each tank
```{r, eval = F, "follicle area plot by tank"}
ggplot(data = female_data, aes(x = tank,y = perc_follicle_area, group = as.factor(tank), fill = pH)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.2) + geom_jitter(shape =16, position= position_jitter(0.1)) + ylab("follicle area (%)") + theme_bw()
```

anova to see if there is a tank or a pH effect
```{r, eval = F}
model <- aov(perc_follicle_area ~ pH * tank, data = female_data)
summary(model)
```


Plot percent egg area for each treatment group
```{r, eval = F,  "egg area plot"}
ggplot(data = female_data, aes(x = pH,y = perc_egg_area, group = as.factor(pH), fill = pH)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.2) + geom_jitter(shape =16, position= position_jitter(0.1)) + ylab("egg area (%)") + theme_bw()
```

Run wilcox test to see if egg area is significantly different
```{r, eval = F}
pH6.8 <- subset(female_data, pH == "6.8", perc_egg_area, drop = TRUE)
pHamb <- subset(female_data, pH == "amb", perc_egg_area, drop = TRUE)
wt <- wilcox.test(pH6.8, pHamb)
print(wt$p.value)
```

Plot percent egg area for each tank
```{r, eval =F, "egg area plot by tank"}
ggplot(data = female_data, aes(x = tank,y = perc_egg_area, group = as.factor(tank), fill = pH)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.2) + geom_jitter(shape =16, position= position_jitter(0.1)) + ylab("egg area (%)") + theme_bw()
```

anova to see if there is a tank or a pH effect
```{r, eval = F}
model <- aov(perc_egg_area ~ pH * tank, data = female_data)
summary(model)
```



Plot egg:follicle ratio for each treatment group
```{r, eval =F, "egg-follicle ratio plot" }
ggplot(data = female_data, aes(x = pH,y = follicle_egg_ratio, group = as.factor(pH), fill = pH)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.2) + geom_jitter(shape =16, position= position_jitter(0.1)) + ylab("egg:follicle ratio") + theme_bw()
```

Run wilcox test to see if egg:follicle ratio is significantly different
```{r,eval =F}
pH6.8 <- subset(female_data, pH == "6.8", follicle_egg_ratio, drop = TRUE)
pHamb <- subset(female_data, pH == "amb", follicle_egg_ratio, drop = TRUE)
wt <- wilcox.test(pH6.8, pHamb)
print(wt$p.value)

```

Plot egg:follicle ratio for each tank
```{r, eval =F, "egg:follicle ratio plot by tank"}
ggplot(data = female_data, aes(x = tank,y = follicle_egg_ratio, group = as.factor(tank), fill = pH)) + geom_violin(trim = FALSE) + geom_boxplot(width = 0.2) + geom_jitter(shape =16, position= position_jitter(0.1)) + ylab("egg:follicle ratio") + theme_bw()
```

anova to see if there is a tank or a pH effect
```{r, eval =F}
model <- aov(follicle_egg_ratio ~ pH * tank, data = female_data)
summary(model)
```