---
title: "Ca Assay Analysis"
author: "Shelly Trigg"
date: "12/7/2020"
output: github_document 
---

load libraries
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
```

load data
```{r}
ca_data <- read_xlsx("../../data/hemolymph_calcium_assay/20201125/CompiledData.xlsx")

sample_info <- read_xlsx("../../data/hemolymph_calcium_assay/20201125/CompiledData.xlsx", sheet = 'SampleInfo' )

histology_data <- read_xlsx("../../data/histology/HistologyScores_STRIGG.xlsx")
```


calculate the avg abs for each standard
```{r}
ca_data_sum <- ca_data %>% group_by(Plate, ID) %>% summarise(mean_abs = mean(`Absorbance @ 570 (1.0s) (A)`))

```

calculate the corrected absorbance by subtracting STD_A from all measurements
```{r}
ca_data_sum_corrected <- ca_data_sum %>% group_by(Plate) %>% mutate(correct_abs = mean_abs - mean_abs[which(ID == "STD_A")])

```

add column with std concentrations
```{r}
ca_data_sum_corrected$conc <- NA

for (i in 1:nrow(ca_data_sum_corrected)){
  if(ca_data_sum_corrected$ID[i]=="STD_A"){
    ca_data_sum_corrected$conc[i] <- 0
  }
  if(ca_data_sum_corrected$ID[i]=="STD_B"){
    ca_data_sum_corrected$conc[i] <- 0.5
  }
  if(ca_data_sum_corrected$ID[i]=="STD_C"){
    ca_data_sum_corrected$conc[i] <- 1
  }
  if(ca_data_sum_corrected$ID[i]=="STD_D"){
    ca_data_sum_corrected$conc[i] <- 2
  }
  if(ca_data_sum_corrected$ID[i]=="STD_E"){
    ca_data_sum_corrected$conc[i] <- 4
  }
  if(ca_data_sum_corrected$ID[i]=="STD_F"){
    ca_data_sum_corrected$conc[i] <- 6
  }
  if(ca_data_sum_corrected$ID[i]=="STD_G"){
    ca_data_sum_corrected$conc[i] <- 8
  }
  if(ca_data_sum_corrected$ID[i]=="STD_H"){
    ca_data_sum_corrected$conc[i] <- 10
  }
}

```

plot corrected STD abs
```{r}

ggplot(ca_data_sum_corrected[grep("STD",ca_data_sum_corrected$ID),], aes(conc,correct_abs)) + geom_point() + facet_wrap(~Plate)
```

run linear models to calculate the slope and intercept
```{r}

#plate 1
summary(lm(conc ~ correct_abs, data = ca_data_sum_corrected[which(ca_data_sum_corrected$Plate ==1 & substr(ca_data_sum_corrected$ID,1,3) == "STD"),]))

#plate 2
summary(lm(conc ~ correct_abs, data = ca_data_sum_corrected[which(ca_data_sum_corrected$Plate ==2 & substr(ca_data_sum_corrected$ID,1,3) == "STD"),]))

#plate 3
summary(lm(conc ~ correct_abs, data = ca_data_sum_corrected[which(ca_data_sum_corrected$Plate ==3 & substr(ca_data_sum_corrected$ID,1,3) == "STD"),]))

#plate 4
summary(lm(conc ~ correct_abs, data = ca_data_sum_corrected[which(ca_data_sum_corrected$Plate ==4 & substr(ca_data_sum_corrected$ID,1,3) == "STD"),]))


# I also did this in excel and I get different values (see below); since I trust excel more than this lm, I'm going with the excel values

#excel slope and intercept values
# plate 1:
  # slope: 0.0999
  # intercept: -0.0064

#plate 2: 
  # slope: 0.1067
  # intercept: -0.0057

#plate 3: 
  # slope: 0.1057
  # intercept: -0.0038

#plate 4:
  # slope: 0.1051
  # intercept: -0.0002
```

calculate calcium concentration from slope
```{r}

# conc. = (corrected sample abs - y intercept / slope) * sample dilution
for (i in 1:nrow(ca_data_sum_corrected)){
  if(substr(ca_data_sum_corrected$ID[i],1,3)!="STD" & ca_data_sum_corrected$Plate[i] == 1){
    ca_data_sum_corrected$conc[i] <- ((ca_data_sum_corrected$correct_abs[i] + 0.0064) / 0.0999) * 4
  }
  if(substr(ca_data_sum_corrected$ID[i],1,3)!="STD" & ca_data_sum_corrected$Plate[i] == 2){
    ca_data_sum_corrected$conc[i] <- ((ca_data_sum_corrected$correct_abs[i] + 0.0057) / 0.1067) * 4
  }
  if(substr(ca_data_sum_corrected$ID[i],1,3)!="STD" & ca_data_sum_corrected$Plate[i] == 3){
    ca_data_sum_corrected$conc[i] <- ((ca_data_sum_corrected$correct_abs[i] + 0.0038) / 0.1057) * 4
  }
  if(substr(ca_data_sum_corrected$ID[i],1,3)!="STD" & ca_data_sum_corrected$Plate[i] == 4){
    ca_data_sum_corrected$conc[i] <- ((ca_data_sum_corrected$correct_abs[i] + 0.0002) / 0.1051) * 4
  }
}

#correct conc for ID 01/23-037L which was a 1:7 dilution (5uL in 35uL total)
ca_data_sum_corrected[which(ca_data_sum_corrected$ID == "01/23-037L"),"conc"] <- (data.frame((ca_data_sum_corrected[which(ca_data_sum_corrected$ID == "01/23-037L"),"correct_abs"]) + 0.0064) / 0.0999) * 7

```

match up with treatment and time info
```{r}
#first check ID is the exact same as the label
head(sample_info[which(sample_info$ID!= sample_info$Label),])

#now create ID column to match with plate reader data
sample_info$ID <- paste0(substr(sample_info$Date,5,6),"/",substr(sample_info$Date,7,8),"-",sample_info$Label)



#merge sample info with plate reader data
ca_data_sum_corrected_info <- merge(ca_data_sum_corrected, sample_info, by = "ID")


#create a column with the number of sampling points called "freq"
ca_data_sum_corrected_info <- merge(ca_data_sum_corrected_info,table(Label = ca_data_sum_corrected_info$Label), by =  "Label")

#create a column with factors to use for labeling individuals with multiple sampling points

ca_data_sum_corrected_info$name <- factor(ifelse(ca_data_sum_corrected_info$Freq !=1,paste0(ca_data_sum_corrected_info$Label,"_",ca_data_sum_corrected_info$Freq),1))

```

plot

```{r}
ggplot(ca_data_sum_corrected_info, aes(x = factor(Date), y = conc, group = interaction(Treatment,factor(Date)),color = Treatment)) + geom_boxplot(outlier.shape = NA) + geom_point(shape = ca_data_sum_corrected_info$name,size=1,position = position_jitterdodge(jitter.width = 0.2)) + theme_bw() + labs(x = "Date", y = "Calcium conc. (mg/dL)") + scale_shape_manual(values=1:nlevels(ca_data_sum_corrected_info$name))  + scale_color_manual(values = c("gray50", "chocolate"))

```

run aov


```{r}
aov_data <- aov(conc ~ Treatment * Date, data = ca_data_sum_corrected_info)
summary(aov_data)

```

Plot data but separate females from males
```{r}
ggplot(ca_data_sum_corrected_info, aes(x = factor(Date), y = conc, group = interaction(Sex,Treatment,factor(Date)),color = Treatment, fill = Sex)) + geom_boxplot(outlier.shape = NA) + geom_point(shape = ca_data_sum_corrected_info$name,size=1,position = position_jitterdodge(jitter.width = 0.1)) + theme_bw() + labs(x = "Date", y = "Calcium conc. (mg/dL)") + scale_shape_manual(values=1:nlevels(ca_data_sum_corrected_info$name)) + scale_color_manual(values = c("gray50", "chocolate")) + scale_fill_manual(values = c("wheat1","lightblue1"))

```

Convert date to days
```{r}
#create a column for days to go into 
ca_data_sum_corrected_info$days <- NA


for(i in 1:nrow(ca_data_sum_corrected_info)){
  if(ca_data_sum_corrected_info$Date[i] == "20181119"){
    ca_data_sum_corrected_info$days[i] <- "7"
  }
  if(ca_data_sum_corrected_info$Date[i] == "20190104"){
    ca_data_sum_corrected_info$days[i] <- "53"
  }
  if(ca_data_sum_corrected_info$Date[i] == "20190123"){
    ca_data_sum_corrected_info$days[i] <- "72"
  }
  if(ca_data_sum_corrected_info$Date[i] == "20190221"){
    ca_data_sum_corrected_info$days[i] <- "93 + 8 day recovery"
  }
}

ca_data_sum_corrected_info$days <- factor(ca_data_sum_corrected_info$days, levels = c("7", "53", "72", "93 + 8 day recovery"))

count_summary <- ca_data_sum_corrected_info %>% group_by(Sex,days) %>% count()

ca_data_sum_corrected_info$Treatment <- gsub("Elevated","low (6.8)",ca_data_sum_corrected_info$Treatment)
ca_data_sum_corrected_info$Treatment <- gsub("Ambient","ambient (7.8)",ca_data_sum_corrected_info$Treatment)

```


Separate plots for males and females
```{r}

#ggplot(ca_data_sum_corrected_info, aes(x = days, y = conc, group = interaction(Sex,Treatment,days),fill = Treatment)) + geom_boxplot(outlier.shape = NA) + geom_point(size=1,shape = 1,position = position_jitterdodge(jitter.width = 0.1)) +  theme_bw() + theme(axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2))  + labs(x = "Days", y = "Calcium conc. (mg/dL)") + facet_wrap(~Sex) 



#define regression line coefficients for female data
coefs_a <- coef(lm(conc ~ days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info$Sex == "F"),]))

coefs_l <- coef(lm(conc ~ days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Treatment == "low (6.8)" & ca_data_sum_corrected_info$Sex == "F"),]))

#define regression line coefficients for male data
coefs_am <- coef(lm(conc ~ days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info$Sex == "M"),]))

coefs_lm <- coef(lm(conc ~ days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Treatment == "low (6.8)" & ca_data_sum_corrected_info$Sex == "M"),]))

#plot female data
f <- ggplot(ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Sex == "F"),], aes(x = days, y = conc, group = interaction(Treatment,days),color = Treatment)) + geom_boxplot(outlier.shape = NA) + geom_point(size=1,shape = 1,position = position_jitterdodge(jitter.width = 0.1)) +  theme_bw() + theme(axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2))  + labs(x = "Days", y = "Calcium conc. (mg/dL)") + scale_color_manual(values = c("#00BFC4", "#F8766D")) + geom_abline(intercept = coefs_a[1], slope = coefs_a[2], col = "#00BFC4") + geom_abline(intercept = coefs_l[1], slope = coefs_l[2], col = "#F8766D") + ggtitle("female")

#plot male data
m <- ggplot(ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Sex == "M"),], aes(x = days, y = conc, group = interaction(Treatment,days),color = Treatment)) + geom_boxplot(outlier.shape = NA) + geom_point(size=1,shape = 1,position = position_jitterdodge(jitter.width = 0.1)) +  theme_bw() + theme(axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2))  + labs(x = "Days", y = "Calcium conc. (mg/dL)") + scale_color_manual(values = c("#00BFC4", "#F8766D")) + geom_abline(intercept = coefs_am[1], slope = coefs_am[2], col = "#00BFC4") + geom_abline(intercept = coefs_lm[1], slope = coefs_lm[2], col = "#F8766D") + ggtitle("male")


jpeg("img/CalciumXtime.jpg", width = 7, height = 4, units = "in", res = 300)
ggarrange(f,m, common.legend = T)
dev.off()
```

run aov for female data
```{r}
f_aov_data <- aov(conc ~ Treatment * days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Sex == "F"),])
summary(f_aov_data)

```

run aov for male data
```{r}
m_aov_data <- aov(conc ~ Treatment * days, data = ca_data_sum_corrected_info[which(ca_data_sum_corrected_info$Sex == "M"),])
summary(m_aov_data)

```


add staging data
```{r}
#create column for Geoduck.ID
ca_data_sum_corrected_info$Geoduck.ID <- gsub("L", "", ca_data_sum_corrected_info$Label)

ca_data_sum_corrected_info_stage <- merge(ca_data_sum_corrected_info,histology_data[,c("Geoduck.ID", "Stage_Trigg")], by = "Geoduck.ID")


ca_data_sum_corrected_info_stage$Treatment <- gsub("Elevated", "low (6.8)", ca_data_sum_corrected_info_stage$Treatment)

ca_data_sum_corrected_info_stage$Treatment <- gsub("Ambient", "ambient (7.8)", ca_data_sum_corrected_info_stage$Treatment)

ca_data_sum_corrected_info_stage$Sex <- gsub("F", "Female", ca_data_sum_corrected_info_stage$Sex)
ca_data_sum_corrected_info_stage$Sex <- gsub("M", "Male", ca_data_sum_corrected_info_stage$Sex)

#remove animal with NA stage
ca_data_sum_corrected_info_stage <- ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Stage_Trigg != "NA"),]

#define regression line coefficients for female data
coefs_a <- coef(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info_stage$Sex == "Female"),]))

coefs_l <- coef(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "low (6.8)" & ca_data_sum_corrected_info_stage$Sex == "Female"),]))

r_a <- summary(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info_stage$Sex == "Female"),]))$r.squared 

r_l <- summary(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "low (6.8)" & ca_data_sum_corrected_info_stage$Sex == "Female"),]))$r.squared 




#define regression line coefficients for male data
coefs_am <- coef(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info_stage$Sex == "Male"),]))

r_am <- round(summary(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "ambient (7.8)" & ca_data_sum_corrected_info_stage$Sex == "Male"),]))$r.squared, digits = 3) 

coefs_lm <- coef(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "low (6.8)" & ca_data_sum_corrected_info_stage$Sex == "Male"),]))

r_lm <- round(summary(lm(conc ~ Stage_Trigg, data = ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Treatment == "low (6.8)" & ca_data_sum_corrected_info_stage$Sex == "Male"),]))$r.squared, digits = 3)



ca_data_sum_corrected_info_stage_m <- ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Sex =="Male"),]

ca_data_sum_corrected_info_stage_f <- ca_data_sum_corrected_info_stage[which(ca_data_sum_corrected_info_stage$Sex =="Female"),]


f <- ggplot(ca_data_sum_corrected_info_stage_f[which(ca_data_sum_corrected_info_stage_f$Date == "20190123" |ca_data_sum_corrected_info_stage_f$Date == "20190221"  ),], aes(x = Stage_Trigg, y = conc, group = interaction(Sex,Treatment,Stage_Trigg),color = Treatment)) +  geom_point(position = position_dodge(width = 0.2),size=1.5, aes(shape = factor(Date))) +  theme_bw() + theme(axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2))  + labs(x = "Stage (based on histology)", y = "Calcium conc. (mg/dL)", shape = "Date", color = "pH")+ scale_color_manual(values = c("#00BFC4", "#F8766D")) + geom_abline(intercept = coefs_a[1], slope = coefs_a[2], col = "#00BFC4") + geom_abline(intercept = coefs_l[1], slope = coefs_l[2], col = "#F8766D") + ggtitle("female")

m <- ggplot(ca_data_sum_corrected_info_stage_m[which(ca_data_sum_corrected_info_stage_m$Date == "20190123" |ca_data_sum_corrected_info_stage_m$Date == "20190221"  ),], aes(x = Stage_Trigg, y = conc, group = interaction(Sex,Treatment,Stage_Trigg),color = Treatment)) +  geom_point(position = position_dodge(width = 0.2),size=1.5, aes(shape = factor(Date))) +  theme_bw() + theme(axis.text.x = element_text(angle=90,hjust=0.95,vjust=0.2))  + labs(x = "Stage (based on histology)", y = "Calcium conc. (mg/dL)", shape = "Date", color = "pH") + scale_color_manual(values = c("#00BFC4", "#F8766D")) + geom_abline(intercept = coefs_am[1], slope = coefs_am[2], col = "#00BFC4") + geom_abline(intercept = coefs_lm[1], slope = coefs_lm[2], col = "#F8766D") + ggtitle("male") + geom_text(x = 1, y = 40, label = paste0(expression(italic(r)^2)," = ",r_am),color ="#00BFC4" , parse = TRUE) + geom_text(x = 1, y = 38, label = paste(expression(italic(r)^2),"=",r_lm),color ="#F8766D")

jpeg("img/CalciumXstage.jpg", width = 7, height = 4, units = "in", res = 300)
ggarrange(f,m, common.legend = T)
dev.off()

```